Exploring dimensions of biodiversity in Japanese ferns

### 2021-05-28 ms/manuscript.Rmd
Add tables for model parameters

```{r setup, include = FALSE}
# Load libraries used for rendering Rmds
library(officedown)
library(officer)
library(flextable)
library(kableExtra)
library(ftExtra)

@@

# Load functions only used when rendering Rmds
source(here::here("R/ms_functions.R"), local = TRUE)

# Run these lines if rendering outside of targets plan
# library(targets)
# library(tarchetypes)
# source(here::here("R/packages.R"))
# source(here::here("R/functions.R"))
```

@@

All biodiversity metrics except for SES of RFD showed a significant relationship with temperature (`r table("env-model-params")`).

@@

```{r env-model-params-prep}
# Format model parameter summary for nice printing as a table
env_model_params_pretty <-
env_model_params %>%
  # Remove intercept
  filter(term != "(Intercept)") %>%
  # Double check that log-likelihood of null model is less than full
  verify(all(loglik_null < loglik_full)) %>%
  select(resp_var, term, estimate, cond_se, t_value, p_value) %>%
  # Format response variable and fixed effect names to look nice
  left_join(resp_var_lookup, by = "resp_var") %>%
  select(-resp_var) %>%
  rename(resp_var = resp_var_pretty) %>%
  left_join(indep_var_lookup, by = "term") %>%
  select(-term) %>%
  rename(term = term_pretty) %>%
  # Round to 2 signif digits
  mutate(across(c(estimate, cond_se, t_value), ~signif(., 2))) %>%
  # Format p-value
  mutate(
    signif = if_else(p_value < 0.05, TRUE, FALSE),
    p_value = case_when(
      p_value < 0.0001 ~ "< 0.0001",
      TRUE ~ signif(p_value, 2) %>% as.character
    )
  ) %>%
  arrange(resp_var, term) %>%
  select(resp_var, term, estimate, cond_se, t_value, p_value, signif)
```

```{r env-model-params-doc, results = "asis", eval = params$doc_type == "doc"}
env_model_params_pretty %>%
  mutate(
    across(where(is.numeric), as.character)) %>%
  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
  as_grouped_data(groups = "resp_var") %>%
  as_flextable(hide_grouplabel = TRUE, col_keys = c("Variable", "Estimate", "Conditional SE", "*t*-value", "*p*-value")) %>%
  bold(i = ~ signif == TRUE, bold = TRUE)
```

@@

>>>>

```{r setup, include = FALSE}
# Run these lines if rendering outside of targets plan
# library(targets)
# library(tarchetypes)
# source(here::here("R/packages.R"))
# source(here::here("R/functions.R"))

# Load libraries used for rendering Rmds
library(officedown)
library(officer)
library(flextable)
library(kableExtra)
library(ftExtra)

@@

# Load functions only used when rendering Rmds
source(here::here("R/ms_functions.R"), local = TRUE)
```

@@

```{r model-stats}
# Verify the number of cases where models had significant fits for precipitation, precipitation seasonality, and temperature
n_env_precip_sig <- 5
n_env_precip_season_sig <- 2
n_env_temp_sig <- 5

env_model_params %>%
  filter(str_detect(term, "Intercept", negate = TRUE)) %>%
  filter(p_value < 0.05) %>%
  count(term) %>%
  verify(.$n[.$term == "precip"] == n_env_precip_sig) %>%
  verify(.$n[.$term == "precip_season"] == n_env_precip_season_sig) %>%
  verify(.$n[.$term == "temp"] == n_env_temp_sig)

assert_that(all.equal(n_env_precip_sig, n_env_temp_sig))

repro_model_params %>%
  filter(str_detect(term, "Intercept", negate = TRUE)) %>%
  filter(p_value < 0.05) %>%
  count(term) %>%
  verify(.$n[.$term == "percent_apo"] == 3)

# Add a column with estimate ± cond_se, then
# split up table for easy referencing in inline chunks
repro_cond_se <-
repro_model_params %>%
  mutate(
    across(c(estimate, cond_se), ~signif(., 2)),
    est_se = glue("{estimate} ± {cond_se}")
) %>%
  split(.$resp_var) %>%
  map(~split(., .$term))

env_cond_se <-
env_model_params %>%
  mutate(
    across(c(estimate, cond_se), ~signif(., 2)),
    est_se = glue("{estimate} ± {cond_se}")
) %>%
  split(.$resp_var) %>%
  map(~split(., .$term))

# Example:
# parameter estimate +/- SE for SES of RPD in reproductive model:
# repro_cond_se$rpd_obs_z$percent_apo$est_se
```

All spatial models were significant relative to null models without any fixed effects. Most biodiversity metrics showed a significant relationship with temperature and/or precipitation (each significant in `r english2(n_env_temp_sig)` out of six models), but precipitation seasonality was only significant in `r english2(n_env_precip_season_sig)` models (`r table("env-model-params")`). Models fit using % apomictic taxa instead of temperature had lower cAIC, indicating a better fit. All biodiversity metrics with a genetic component (PE *p*-value, SES of PD, SES of RPD) each had a significant, negative relationship to % apomictic taxa (`r table("repro-model-params")`).

@@

```{r model-params-prep}
# Format model parameter summaries for nice printing as a table
model_params_pretty <-
  # Combine environmental and reproductive model results to avoid repitition
  bind_rows(env_model_params, repro_model_params) %>%
  # Remove intercept
  filter(term != "(Intercept)") %>%
  # Double check that log-likelihood of null model is less than full
  verify(all(loglik_null < loglik_full)) %>%
  # Format response variable and fixed effect names to look nice
  left_join(resp_var_lookup, by = "resp_var") %>%
  select(-resp_var) %>%
  rename(resp_var = resp_var_pretty) %>%
  left_join(indep_var_lookup, by = "term") %>%
  select(-term) %>%
  rename(term = term_pretty) %>%
  # Round to 2 signif digits
  mutate(across(c(estimate, cond_se, t_value), ~signif(., 2))) %>%
  # Format p-value
  mutate(
    signif = if_else(p_value < 0.05, TRUE, FALSE),
    p_value = case_when(
      p_value < 0.0001 ~ "< 0.0001",
      TRUE ~ signif(p_value, 2) %>% as.character
    )
  ) %>%
  arrange(data_type, resp_var, term) %>%
  select(resp_var, term, estimate, cond_se, t_value, p_value, signif, data_type)

env_model_params_pretty <- model_params_pretty %>%
  filter(data_type == "env") %>%
  select(-data_type)

repro_model_params_pretty <- model_params_pretty %>%
  filter(data_type == "repro") %>%
  select(-data_type)
```

```{r env-model-params-doc, results = "asis", eval = params$doc_type == "doc"}
# Print out environmental parameters formatted for word document
env_model_params_pretty %>%
  mutate(
    across(where(is.numeric), as.character)) %>%
  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
  as_grouped_data(groups = "resp_var") %>%
  as_flextable(hide_grouplabel = TRUE, col_keys = c("Variable", "Estimate", "Conditional SE", "*t*-value", "*p*-value")) %>%
  bold(i = ~ signif == TRUE, bold = TRUE) %>%
  width(1, width = 1.5) %>%
  width(2:5, width = 1)
```

@@

```{r repro-model-params-doc, results = "asis", eval = params$doc_type == "doc"}
# Print out environmental parameters formatted for word document
repro_model_params_pretty %>%
  mutate(
    across(where(is.numeric), as.character)) %>%
  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
  as_grouped_data(groups = "resp_var") %>%
  as_flextable(hide_grouplabel = TRUE, col_keys = c("Variable", "Estimate", "Conditional SE", "*t*-value", "*p*-value")) %>%
  bold(i = ~ signif == TRUE, bold = TRUE) %>%
  width(1, width = 1.5) %>%
  width(2:5, width = 1)
```

```{r repro-model-params-pdf, results = "asis", eval = params$doc_type == "pdf"}
# Define row groups
kbl_groups <-
  repro_model_params_pretty %>%
  transmute(resp_var, row_num = 1:nrow(.)) %>%
  group_by(resp_var) %>%
  summarize(start = min(row_num), end = max(row_num), .groups = "drop") %>%
  split(.$resp_var)

signif_rows <- repro_model_params_pretty %>% pull(signif) %>% which()

repro_model_params_pretty %>%
  mutate(
    across(where(is.numeric), as.character)) %>%
  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
  select(-resp_var, -signif) %>%
  kbl(booktabs = TRUE) %>%
  kable_styling() %>%
  pack_rows(kbl_groups[[1]]$resp_var, kbl_groups[[1]]$start, kbl_groups[[1]]$end) %>%
  pack_rows(kbl_groups[[2]]$resp_var, kbl_groups[[2]]$start, kbl_groups[[2]]$end) %>%
  pack_rows(kbl_groups[[3]]$resp_var, kbl_groups[[3]]$start, kbl_groups[[3]]$end) %>%
  row_spec(row = signif_rows, bold = TRUE)
```

**`r table("repro-model-params")`**. Parameter estimates of the effect of reproductive mode and environmental variables (precipitation) on biodiversity metrics in the ferns of Japan. Significant effects (*p* < 0.05) in bold; significance assessed with likelihood ratio tests (see Methods). All models include random effect of geographical coordinates.
